---
title: "Data Processing and Exploratory analyses"
author: 
  - "Jake W. Coldiron,"
  - "Zixuan Qiu,"
  - "Tianyuan Deng,"
  - "Aung Nay Win,"
  - "Yuqi Cheng"
output: github_document
---

**Date:** 2023-11-10

```{r message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(plotly)
library(broom)
library(modelr)
```

# Data processing

```{r message=FALSE}
#input the New York HIV data from NYC opendata 
nyc_HIV1=
  GET("https://data.cityofnewyork.us/resource/ykvb-493p.csv",
  query = list("$limit" = 10000)) |>
  content("parsed")|>
  filter(year >= 2017)


nyc_HIV2=
  GET("https://data.cityofnewyork.us/resource/dxnu-p2qd.csv",
  query = list("$limit" = 22000)) |>
  content("parsed")|>
  filter(year >= 2017)
  
```

```{r message=FALSE}
#HIV service resource 
HIV_service=
  GET("https://data.cityofnewyork.us/resource/pwts-g83w.csv",
  query = list("$limit" = 1000)) |>
  content("parsed")|>
  select(unique_id,facilityname,service_type, address,borough,state,zipcode,latitude,longitude)
  
```
**A short  clear description of where and how data were obtained. explain the variables means in the data**
**Calculate the sum, mean, median, standard deviation of HIV/AIDS indicators** 


# Visualizations

**Shows the number of HIV/AIDS diagnoses by race, sex, and age group.**
```{r}
#work
#work
colnames(nyc_HIV1)
colnames(nyc_HIV2)
head(nyc_HIV1)
library(ggplot2)
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

Female <- nyc_HIV1 %>% 
  filter(sex == "Female" & borough == "All" & race !="All")

Male <- nyc_HIV1 %>% 
  filter(sex == "Male"  & borough == "All" & race !="All")

graph1 = bind_rows(Female, Male)

nyc_HIV1$hiv_diagnoses_num <- as.integer(nyc_HIV1$hiv_diagnoses_num)
nyc_HIV2$hiv_diagnoses_num <- as.integer(nyc_HIV2$hiv_diagnoses_num)

ggplot(graph1, aes(x=race, y=hiv_diagnoses_num, fill=sex)) +
  geom_bar(stat="identity", position=position_dodge()) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Race", y = "Number of HIV/AIDS Diagnoses", fill = "Sex")

# Graph Age 
age_summary <- nyc_HIV2 %>%
  filter(age != "All" & race == "All" & borough == "All") %>%
  group_by(age) %>%
  summarise(total_hiv_diagnoses = sum(hiv_diagnoses_num))

ggplot(age_summary, aes(x = age, y = total_hiv_diagnoses)) +
  geom_bar(stat = "identity", color = "yellow", width = 0.7) +
  labs(title = "Total HIV Diagnoses by Age Group",
       x = "Age Group",
       y = "Total HIV Diagnoses") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  geom_text(aes(label=total_hiv_diagnoses), vjust=-0.5, color="black", size=3.5)
```
summary:
Both the nyc_HIV1 dataset and the nyc_HIV1 data have 11 variables, nyc_HIV1 is the data related to the number of patients with HIV by gender and nyc_HIV2 is the data related to the number of patients with HIV by age. When analyzing the number of HIV patients by race, the results show that Latino/Hispanic races are used for high prevalence and there are more males than females. When analyzing the number of HIV patients by age group, the results showed that 20-29 year olds had the highest number of total diagnoses and 13-19 year olds had the lowest number of diagnoses.

**Show the map of NYC HIV service resources location.**
```{r}
if (!require(leaflet)) install.packages('leaflet')
library(leaflet)
filtered_data = HIV_service[HIV_service$longitude != 0 & HIV_service$latitude != 0,]
# Create the map
m <- leaflet(filtered_data) %>%
  addProviderTiles(providers$OpenStreetMap) %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    popup = ~paste("<strong>Facility Name:</strong>", facilityname,
                   "<br><strong>Service Type:</strong>", service_type,
                   "<br><strong>Address:</strong>", address,
                   "<br><strong>Borough:</strong>", borough,
                   "<br><strong>Zip:</strong>", zipcode)
  )

# Print the map
print(m)
```
summary:
Here, an interactive map was created using the leaflet package to visualize the distribution of HIV service facilities and detailed information to help better understand the allocation of resources for HIV prevention and treatment.

**Show the distribution of the HIV/AIDS diagnoses by neighborhood.** 
```{r}
#work
# I'll see if I can find the log&lat for the neighborhoods
diagnoses <- nyc_HIV1 %>% 
  select(borough, hiv_diagnoses_num)

# Plot bar chart
ggplot(diagnoses, aes(x = reorder(borough, hiv_diagnoses_num), y = hiv_diagnoses_num)) +
  geom_col() +
  labs(x = "Neighborhood",
       y = "Number of HIV Diagnoses",
       title = "HIV Diagnoses by Neighborhood") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```
```{r}
filtered_data = HIV_service[HIV_service$longitude != 0 & HIV_service$latitude != 0,]
# Create the map
m <- leaflet(filtered_data) %>%
  addTiles("https://server.arcgisonline.com/ArcGIS/rest/services/World_Street_Map/MapServer/tile/{z}/{y}/{x}") %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    radius = 4,
    popup = ~paste("<strong>Facility Name:</strong>", facilityname,
                   "<br><strong>Service Type:</strong>", service_type,
                   "<br><strong>Address:</strong>", address,
                   "<br><strong>Borough:</strong>", borough,
                   "<br><strong>Zip:</strong>", zipcode)
  )
m <- m %>% addCircleMarkers(
  radius = sqrt(nyc_HIV1$hiv_diagnoses_num),
  fillOpacity = 0.4,
  color = "orange",
  stroke = TRUE,
  label = sprintf("HIV Diagnoses: %s", nyc_HIV1$hiv_diagnoses_num),
  labelOptions = labelOptions(
    textOnly = TRUE,
    direction = "auto"
  )
)
m
```
summary:
The number of HIV cases varies by community, with the Staten Island community having a much smaller number of confirmed cases, the Brooklyn community having a larger number of confirmed cases, and the other communities having a similar number of cases.


**Trend analysis:analyze the trend of the number of HIV/AIDS diagnoses over time.**
```{r}
#work
All <- nyc_HIV1 %>% 
  filter(sex == "All" & race == "All" & borough == "All")

# Plot diagnoses over time
ggplot(All, aes(x = year, y = hiv_diagnoses_num)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "HIV Diagnoses Trend (All Sex)",
    x = "Year",
    y = "Number of Diagnoses"
  )
```
summary:
The number of patients diagnosed in different years showed significant differences. The overall trend is decreasing from year to year. This also reflects the growing importance that has been placed on HIV.



# Exploratory statistical analyses.
**Correlation analysis:After the Visualizations is there any association between different variables such as neighborhood, age, gender, and race.**
**T-test for HPV prevalence in gender-specific populations.**
```{r}
female_data <- nyc_HIV1[nyc_HIV1$sex == "Female",]$hiv_diagnoses_num
male_data <- nyc_HIV1[nyc_HIV1$sex == "Male",]$hiv_diagnoses_num
# Perform T test
t.test(female_data, male_data)
```
summary:
The t-test passed the test of significance indicating that there was a significant difference in the number of males and females with HIV.

**Test whether "hiv_diagnoses_num" is significantly different across ethnicities.**
```{r}
anova1 <- anova(lm(hiv_diagnoses_num ~ race, data = nyc_HIV1))
print(anova1)
```
summary:
The ANOVA passed the test of significance, indicating that there is indeed a significant difference in the number of people of different races who have HIV.

**Test whether "hiv_diagnoses_num" is significantly different between age groups.**
```{r}
anova2 <- anova(lm(hiv_diagnoses_num ~ age, data = nyc_HIV2))
print(anova2)
```
summary:
The ANOVA passed the test of significance indicating that there is a significant difference in the number of HIV in different age groups.

**Test whether there is a significant difference between people of different races and different age groups.**
```{r}
race_table <- table(nyc_HIV1$race, nyc_HIV1$sex)
race_table
# Perform chi-square test
chisq.test(race_table)
```
summary:
The chi-square test passes the test of significance, indicating that there is a significant relationship between race and gender.

# linear regression analysis
```{r}
# Defining a Unique Thermal Coding Processor
library(caret)
encoder <- dummyVars(~ borough + sex + race, data = nyc_HIV1)
# Solo thermal coding of data
encoded_data <- predict(encoder, newdata = nyc_HIV1)
# Merging the coded and original data sets
nyc_HIV3 <- cbind(nyc_HIV1, encoded_data)
nyc_HIV4<-nyc_HIV3[,-c(2,3,4,5,7,8,9,10,11,12,18,21)]
model <- lm(hiv_diagnoses_num ~. , data = nyc_HIV4)
summary(model)
#Drawing demonstration
par(mfrow=c(2, 2))  
plot(model)
# Plot of predicted values against residuals
nyc_HIV4=nyc_HIV4|>
  add_predictions(model, var = "fitted_values")|>
  add_residuals(model, var = "residuals")

ggplot(nyc_HIV4, aes(x = fitted_values, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals")
```


