---
title: "Data Processing and Exploratory analyses"
author: 
  - "Jake W. Coldiron,"
  - "Zixuan Qiu,"
  - "Tianyuan Deng,"
  - "Aung Nay Win,"
  - "Yuqi Cheng"
output: github_document
---

**Date:** 2023-11-10

```{r message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(plotly)
```

# Data processing

```{r message=FALSE}
#input the New York HIV data from NYC opendata 
nyc_HIV1=
  GET("https://data.cityofnewyork.us/resource/ykvb-493p.csv",
  query = list("$limit" = 10000)) |>
  content("parsed")|>
  filter(year >= 2017)


nyc_HIV2=
  GET("https://data.cityofnewyork.us/resource/dxnu-p2qd.csv",
  query = list("$limit" = 22000)) |>
  content("parsed")|>
  filter(year >= 2017)
  
```

```{r message=FALSE}
#HIV service resource 
HIV_service=
  GET("https://data.cityofnewyork.us/resource/pwts-g83w.csv",
  query = list("$limit" = 1000)) |>
  content("parsed")|>
  select(unique_id,facilityname,service_type, address,borough,state,zipcode,latitude,longitude)
  
```
**A short  clear description of where and how data were obtained. explain the variables means in the data**
**Calculate the sum, mean, median, standard deviation of HIV/AIDS indicators** 


# Visualizations

**Shows the number of HIV/AIDS diagnoses by race, sex, and age group.**
```{r}
#work
```
summary:
**Show the map of NYC HIV service resources location.**
```{r}
#work
```
summary:

**Show the distribution of the HIV/AIDS diagnoses by neighborhood.** 
```{r}
#work
```
summary:

**Trend analysis:analyze the trend of the number of HIV/AIDS diagnoses over time.**
```{r}
#work
```
summary:

# Exploratory statistical analyses.
**Correlation analysis:After the Visualizations is there any association between different variables such as neighborhood, age, gender, and race.**

```{r}
#work
```
T-test /ANOVA: Compare differences in HIV between two or more groups ( gender or racial).
Chi-square test: Analyze associations between categorical variables (gender, race).
